\chapter{The Network Machine Learning Landscape}
\label{sec:ch1}

In this section, we'll cover the basics of the network machine learning landscape. We learn a lot of high level elements about networks, and introduce some of the basic terminology for network data. This will give you an understanding of the basics of the network data structure, which will make Chapter \ref{sec:ch4} more digestible when we formally introduce network data structures and how you can manipulate them. 

We take a look at the different types of problems you might come across in network machine learning, and how network data fits into the types of problems you want to address. Then, we'll see how a particular branch of mathematics,  \emph{statistics}, fits into the machine learning landscape. This chapter will be very high level, but will allow you to see how the pieces of the book fit together in the grand scheme of things. 

Here are some basic questions that we will approach in these next sections:
\begin{enumerate}
    \item How do networks fit into the world of machine learning, and why are they important?
    \item What does it mean to learn from a network, and what kinds of things would you be trying to learn?
    \item How can we use network-valued data to understand the world better?
\end{enumerate}

Having a good understanding of the high level will help you see how the lower level parts fit together in terms of your ultimate goal of applying network machine learning to new and exciting problems in your work. 

\newpage 

\input{foundations/ch1/what-is-a-network}
% \input{foundations/ch1/why-study-networks}
\input{foundations/ch1/types-nml-problems}
\input{foundations/ch1/challenges-of-nml}


\bibliographystyle{vancouver}
\bibliography{references}