\chapter{End-to-end Biology Network Machine Learning Project}
\label{sec:ch2}

In this chapter, we try to motivate network machine learning with some behind the scenes work, to give you a feel for what the process is really like from start to finish. While some of the material in this section might seem a little bit opaque at first, we try to keep the details lighter and the big picture in focus. You can jump around with network machine learning in a real setting, before we start jumping right into the nitty, gritty details. This will give you a chance to ensure that you have all of the dependencies set up appropriately from the beginning. Things run a lot smoother, in our experience, when you start with a working programming environment. We have the following sections:
\begin{enumerate}
    \item Section \ref{sec:ch2:bigpicture} discusses how to approach a network machine learning problem.
    \item Section \ref{sec:ch2:getdata} establishes how to set up your local environment to use this textbook, and how to grab some example data. 
    \item Section \ref{sec:ch2:prepare} details how to prepare your data for downstream analysis.
    \item Section \ref{sec:ch2:select} covers the algorithm selection process for network data.
    \item Section \ref{sec:ch2:finetune} demonstrates the tuning process for network machine learning algorithms.
    \item Section \ref{sec:ch2:discover} illustrates how we can use network machine learning to uncover new insights about network data.
\end{enumerate}

\newpage 

\input{foundations/ch2/big-picture}
\input{foundations/ch2/get-the-data}
\input{foundations/ch2/prepare-the-data}
\input{foundations/ch2/select-and-train}
\input{foundations/ch2/fine-tune}
\input{foundations/ch2/discover-and-visualize}

\bibliographystyle{vancouver}
\bibliography{references}